---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}

deaths <- av %>% pivot_longer(cols = c("Death1", "Death2", "Death3", "Death4", "Death5"),  
                              names_to = "Time", values_to = "Death")

deaths$Time <- parse_number(deaths$Time)
```

Similarly, deal with the returns of characters.

```{r}
deaths2 <- av %>% pivot_longer(cols = c("Return1", "Return2", "Return3", "Return4", "Return5"), names_to = "Time", values_to = "Return")

deaths$Return <- deaths2$Return

deaths = subset(deaths, select = -c(Return1, Return2, Return3, Return4, Return5))
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
numDeaths = length(which(deaths$Death=="YES"))
numAvengers = length(unique(deaths$Name.Alias))
averageDeaths = numDeaths / numAvengers
averageDeaths
```

The average number of deaths an avenger suffers is 0.546 deaths.

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

------------------------------------------------------------------------------------------------------------------------------------------

Ryan Freidhoff: 

> "Given the Avengers’ 53 years in operation and overall mortality rate, fans of the comics can expect one current or former member to die every seven months or so, with a permanent death occurring once every 20 months."

```{r}
numDeaths = length(which(deaths$Death=="YES"))
timeBetweenDeaths = (53 * 12) / numDeaths
numPermanentDeaths = length(which(deaths$Return=="NO"))
timeBetweenPermanentDeaths = (53 * 12) / numPermanentDeaths
timeBetweenPermanentDeaths
timeBetweenDeaths
```


The average time between deaths in months is about 7.14 months, which strengthens the given statement that someone dies every 7 months or so. The average time between permanent deaths was found to be about 19.85 months, which also backs up the statement that there is a permanent death about every 20 months or so.

------------------------------------------------------------------------------------------------------------------------------------------

Ryan Rockey: 

> "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife..."

```{r}
# Number of times an avenger comes back from their first death
numTime1Returns = length(which(deaths$Return=="YES" & deaths$Time=="1"))

# Number of times an avenger dies
numTime1Deaths = length(which(deaths$Death=="YES" & deaths$Time=="1"))
numTime1Returns / numTime1Deaths
```

> "but only a 50 percent chance they recovered from a second or third death."

```{r}
# Number of times an avenger comes back from their second or third death
numTime23Returns = length(which(deaths$Return=="YES" & (deaths$Time=="2" | deaths$Time=="3")))

# Number of times an avenger dies for the second or third time
numTime23Deaths = length(which(deaths$Death=="YES" & (deaths$Time=="2" | deaths$Time=="3")))

numTime23Returns / numTime23Deaths
```

Doing the analysis shows that avengers have an exactly 2 in 3 chance to come back after their first death and similarly an exactly 1 in 2 chance to come back after their second or third death.

------------------------------------------------------------------------------------------------------------------------------------------
 
Mohamed Bashier:

> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team."

```{r}
avengers <- av %>%
  mutate(Total_Deaths = rowSums(select(., starts_with("Death")) == "YES"))

died_at_least_once <- sum(avengers$Total_Deaths > 0)

died_at_least_once

```
That statement has been fact checked, out of 173 listed avengers, 69 had died at least one time after they joined the team. 

------------------------------------------------------------------------------------------------------------------------------------------

Group member 4: 

> "I counted 89 total deaths — some unlucky Avengers7 are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback."
> "Number of Deaths and Returns"

```{r}
# Calculate the total number of deaths and returns
totalDeaths <- sum(deaths$Death == "YES")
totalReturns <- sum(deaths$Return == "YES")

totalDeaths
totalReturns


```

According to the analysis, there were 89 deaths among the Avengers, and 57 of those deaths resulted in the person's return. This suggests that among the Avengers, there were more deaths than returns.
